---
title: "Data Visualisation with leaflet and plotly"
author: "christopher barkhuizen"
date: "02/07/2019"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

```

## The Data

This analysis uses the City of Cape Town street people dataset. 
The data can be found [here](https://web1.capetown.gov.za/web1/OpenDataPortal/DatasetDetail?DatasetName=Street%20people)

The dataset is based on interviews conducted with homeless people and captures various demographics relating to their age, race, occupation etc 


```{r read_data, cache = TRUE}
library(xlsx)
library(dplyr)
library(leaflet)
library(plotly)

streets <- read.xlsx("./Street people head count 19.8.2015.xlsx", 2)
streets_desc <- read.xlsx("./Street people head count 19.8.2015.xlsx",3)
```

## EDA   

```{r eda, cache = TRUE}

str(streets_desc)

p1 <- plot_ly(x = streets_desc$Race, type = "histogram")
p2 <- plot_ly(x = streets_desc$Age, type = "histogram")
p3 <- plot_ly(x = streets_desc$Gender, type = "histogram")
subplot(p1, p2, p3) %>% layout(showlegend = FALSE)
```


## Street People by Location

First we will show the location of all the people encountered in the survey.
The people are grouped into clusters to give a high level view which can then be drilled down. 
A rectangle is drawn to show the geographic area covered by the survey. 

```{r analyse_and_plot, cache = TRUE}

streets_count <- streets %>% mutate(locn_x_fact = as.factor(streets$X...............GPS.Co.ordinate), locn_y_fact = as.factor(streets$Y..................GPS.Co.ordinate)) %>%  
    group_by(locn_x_fact, locn_y_fact) %>% 
    summarise(count = n()) %>% 
    mutate(lat = as.numeric(as.character(locn_y_fact)), lng = as.numeric(as.character(locn_x_fact)))

# Find the limits of the region for the data analysis 

max_x <- max(streets$X...............GPS.Co.ordinate)
min_x <- min(streets$X...............GPS.Co.ordinate)
max_y <- max(streets$Y..................GPS.Co.ordinate)
min_y <- min(streets$Y..................GPS.Co.ordinate)

streets_count %>% leaflet() %>% addTiles() %>% addMarkers(clusterOptions = markerClusterOptions()) %>% addRectangles(color = "light blue", lat1 = min_y, lat2 = max_y, lng1 = min_x, lng2 = max_x)



```

## Street People by Gender   

The next map shows the gender and number of street people per location with the group size indicated by circle radius.
Two different colours are used to distinguish between male and females.

``` {r location_by_gender, cache = TRUE}

streets_gender <- streets %>% mutate(locn_x_fact = as.factor(streets$X...............GPS.Co.ordinate), locn_y_fact = as.factor(streets$Y..................GPS.Co.ordinate)) %>%  
    group_by(locn_x_fact, locn_y_fact, Gender) %>% 
    summarise(count = n()) %>% 
    mutate(lat = as.numeric(as.character(locn_y_fact)), lng = as.numeric(as.character(locn_x_fact)))


colors <- c("red","blue")
get_colors <- function(i) {return(colors[i])}
sc <- sapply(streets_gender$Gender, get_colors)

streets_gender %>% leaflet() %>% addTiles() %>% addCircles(color = sc, weight = 2, radius = (streets_gender$count * 10)) %>% addLegend(labels = c("Male", "Female"), colors = c("red","blue"))

```

